---
title: SAR model
author: Masatoshi Katabuchi
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
#rmdformats::material
  html_document:
    fig_caption: yes
    theme: spacelab #readable #sandstone #spacelab #flatly
    highlight: pygments #tango #kate
    toc: TRUE
    toc_depth: 2
    number_sections: TRUE
    toc_float:
      collapsed: TRUE
      smooth_scroll: TRUE
      toc: true
---


```{r, echo = F}
rm(list = ls())
```

- nested dataframe `n_d3` and dataframe `sar_dat` contain SAR corrected values (`transformed_metric`)


# FRi, MPD, MNND (MNTD)

- First, I calucated FD metrics.
- We can change the distance matrix from `gower` to `euclidean`.
- I didn't include Functinal Richnees for now.

```{r, cache = TRUE}
library(cluster)
library(FD)

load("./data/SE_dat.rda")

site_name <- paste(rep(0:11, each = 32), rep(7:38, 12), sep = "_")
site_dat <- tibble(site = site_name)
samp_vec <- c("d1", "d2", "d3")

list_fun <- function(metric = c("mpd", "mntd")){
  year_tmp <- c(1976, 1996, 2006)
  mpd_list <- list()
  for (j in 1:3) {
    mpd_mat <- NULL
    mntd_mat <- NULL
    samp <- get(samp_vec[j])
    trait_edit <- trait[colnames(samp), ]
    for (i in 1:8) {
      t_dis <- daisy(trait_edit[, i] %>% 
                     as.matrix,
                   metric = "gower") %>% as.matrix
      colnames(t_dis) <- rownames(t_dis) <- rownames(trait_edit)
      if (metric == "mpd") {
      tmp1 <- mpd(samp, t_dis, abundance.weighte = T)
      } else if (metric == "mntd"){
      tmp1 <- mntd(samp, t_dis, abundance.weighte = T)
      }
      mpd_mat <- cbind(mpd_mat, tmp1)
    }
    colnames(mpd_mat) <- colnames(trait_edit)
    mpd_list[[j]] <- mpd_mat %>%
      as_tibble %>%
      mutate(site = rownames(samp)) %>%
      full_join(., site_dat, by = "site") %>%
      mutate(metric_name = metric) %>%
      mutate(time = year_tmp[j])

  }
  mpd_list
}


mpd_list <- list_fun("mpd")
mntd_list <- list_fun("mntd")

fd_dat <- bind_rows(mpd_list[[1]], 
          mpd_list[[2]],
          mpd_list[[3]],
          mntd_list[[1]],
          mntd_list[[2]],
          mntd_list[[3]]) %>%
  mutate(gx = strsplit(site, "_") %>% sapply("[", 1) %>% as.numeric) %>%
  mutate(gy = strsplit(site, "_") %>% sapply("[", 2) %>% as.numeric) %>%
  mutate(year_site = paste(time, site, sep = "_"))

```

# Merge FD and CWM data

- Three quadrats are missing

```{r}
d1 <- trait_com2 %>%
  filter(time == 1976)

ggplot(d1, aes(x = gx, y = gy, fill = temp)) +
  geom_raster() +
  coord_fixed()

```

- We need to add rows for those missing data

```{r}

trait_com3 <- trait_com2 %>%
  mutate(metric_name = "CWM") %>%
  dplyr::select(1:8,
                site,
                metric_name,
                time,
                gx,
                gy,
                year_site)

gy_tmp <- c(31, 34, 37)
year_tmp <- c(1976, 1996, 2006)
na_dat <- NULL

for (i in 1:3) {
  for (j in 1:3) {
    tmp <- trait_com3 %>%
      slice(1)
    tmp[1:8] <- NA
    tmp[1:8] <- as.numeric(tmp[1:8])
    tmp$gx <- 0
    tmp$gy <- gy_tmp[i]
    tmp$site <- paste(tmp$gx, tmp$gy, sep = "_")
    tmp$time <- year_tmp[j]
    tmp$year_site <- paste(tmp$time, tmp$site, sep = "_")
    na_dat <- rbind(na_dat, tmp)
  }
}

trait_com4 <- bind_rows(na_dat, trait_com3) %>%
  bind_rows(., fd_dat) %>%
  arrange(year_site) %>%
  arrange(metric_name) %>%
  mutate(temp2 = rep(1:(nrow(.)/9), 9) %>% as.character)

n_d <- trait_com4 %>%
  gather(trait_name, val, 1:8) %>%
  group_by(trait_name, metric_name, time) %>%
  nest()

```

# SAR model

```{r}

library(spdep)
library(spatialreg)


d_nb <- cell2nb(12, 32, type = "queen")
d_W <- nb2listw(d_nb, zero.policy = T, style = "W")

sar_fun <- function(data) {
  metric <- data$val
  res <- spautolm(metric ~ 1, listw = d_W, na.action = na.omit)
  #res$fit$fitted.values
  res2 <- res$fit$fitted.values
  res_data <- tibble(temp2 = names(res2), transformed_metric = res2)
  res_data
}


```

We apply the function to the nested data.

```{r, cache = TRUE}

before <- proc.time()
n_d2 <- n_d %>%
  mutate(sar_val = map(data, sar_fun))
after <- proc.time()
after - before


mer_fun <- function(data, sar_val){
  full_join(data, sar_val, by = "temp2")
}


n_d3 <- n_d2 %>%
  mutate(data2 = map2(data, sar_val, mer_fun)) %>%
  dplyr::select(trait_name, metric_name, time, data2)

```

Now we have SAR transformed values for each trait, each trait metric and each census.

```{r}

sar_dat <- n_d3 %>%
  unnest()

n_d3

```

For example, `transfomred_metric` in the table below is SAR transformed CWM SEED.

```{r}

n_d3$data2[[1]] %>%
  DT::datatable()

```

CWM SEED (without SAR)

```{r}
moge <- n_d3$data2[[1]]

ggplot(moge, aes(x = gx, y = gy, fill = val)) +
  geom_raster() +
  coord_fixed()

```

CWM SEED (SAR corrected)

```{r}

ggplot(moge, aes(x = gx, y = gy, fill = transformed_metric)) +
  geom_raster() +
  coord_fixed()

```

```{r}
save.image("./data/SAR_dat.rda")
```

# Computing Environment

```{r}
devtools::session_info()
```
